The Inner Mass-Ratio Distribution of Intermediate-Mass Stars

### 2016-02-12 paper/ms.tex
Switched from gaussian to beta distribution fit, among other small things.

\begin{figure}
\includegraphics[width=\columnwidth]{Imaging_Data.pdf}
\caption{Detection images for all stars in which we detect a companion in the follow-up NIRI data.}
\label{fig:images}
\end{figure}

@@

\begin{equation}
Q(T_\mathrm{eff}) = \sum_k Q(T, v_k) v_k 
\end{equation} 
where $Q(T, v\sin{i})$ is the surface plotted in Figure \ref{fig:detrate_2d} and $v_k$ are the samples from the distribution of $v\sin{i}$. For $T < 6000$ K, we sample $v\sin{i}$ using the gyrochronology relation given in \citet{Barnes2010b}:

@@

We show the marginalized detection rate and mean value of $v\sin{i}$ as a function of temperature in Figure \ref{fig:marginalized}. Both the detection rate and the average $v\sin{i}$ are smoothly varying, and show the expected behaviour with temperature. The detection rate falls with hotter temperatures because the companions are expected to be fast rotators, which are more difficult to detect. ADAM AND SALLY: IS THIS FIGURE NECESSARY OR IS THE TEXT ENOUGH?

@@

Where did HIP 118243 go? It has an imaging detection (though at very wide separation), but nothing in the Companion table.

@@

In the above equation, $\{\vec{x_k}\}$ denotes the full spectrum of star $k$, and $\vec{\theta}$ denotes the parameters for the model we are fitting. $K=360$ is the total sample size for our survey, $Q(q)$ is the completeness function shown in Figure \ref{fig:completeness}, and $P(q|\vec{\theta})$ is the likelihood function for the mass ratio given the model parameters. We fit the data to two distinct distributions: a histogram ($P_1$) and a gaussian distribution ($P_2$):

\begin{align}
 P_1(q|\vec{\theta}) &= \begin{cases}
      \hfill \theta_1 \hfill & q \in \Delta_1 \\
      \hfill \theta_2 \hfill & q \in \Delta_2 \\
      \hfill \ldots \\
      \hfill \theta_7 \hfill & q \in \Delta_7
     \end{cases} \label{eqn:P1} \\
 P_2(q|\vec{\theta}) &= \frac{A}{\sqrt{2\pi\sigma^2}} e^{-\frac{(q - \mu)^2}{2\sigma^2}} \label{eqn:P2} \\
 P_3(q|\vec{\theta}) &= (1-\gamma)q^{-\gamma} \label{eqn:P3}
\end{align}
The constant $A$ in the gaussian distribution is a renormalization factor such that the distribution is only defined from $0 < q < 1$:

\begin{equation}
A = \frac{2}{\erf\left(\frac{\mu}{\sigma \sqrt{2}}\right) - \erf\left(\frac{\mu - 1}{\sigma \sqrt{2}}\right)}
\end{equation}

@@

The gaussian fit only has two parameters ($\mu, \sigma$), and was chosen because it has a similar shape to the histogram resulting from the first model. We use uniform priors on both $\mu$ and $\sigma$, resulting in the posterior probability

\begin{equation}
P_2(\mu, \sigma | \{\vec{x_k}\}) \propto P_2(q|\mu, \sigma) 
\end{equation}

@@

We show the resulting fits in Figure \ref{fig:mrd}. The $1\sigma$ uncertainties in the bin heights from the histogram model are shown as error bars, and we overplot 500 samples of the gaussian fit to show the spread allowed by the data. The best-fit power law is plotted with a green dot-dashed line, as well as the mass ratio distribution expected from random pairing from the Kroupa IMF in yellow.

@@

\begin{figure}
\includegraphics[width=\columnwidth]{MRD_total.pdf}
\caption{Mass ratio distribution for our sample. The data was fit to a histogram, a gaussian distribution, and a power law. The histogram is shown in the solid blue blocks, with $1 \sigma$ uncertainties marked with error bars. The variance of the gaussian fit is shown with 500 samples from the posterior probability distribution for the parameters. We also show the best-fit power law and the mass ratio distribution from random pairing of the Kroupa initial mass function (IMF).}
\label{fig:mrd}
\end{figure}

@@

The most striking feature of the mass ratio distribution shown in Figure \ref{fig:mrd} is the turnover at intermediate q, which the gaussian fit finds occurs at $q = 0.36 \pm 0.04$. Although the power law fit and the Kroupa IMF are visually very poor fits to the data, we formally compare the models to ensure that the different form is not an artifact of binning or simple noise. We make the comparison using the posterior odds:

@@

The integrals are estimated as part of the nested sampling algorithm in the MultiNest code. The odds ratio comparing the gaussian to power law fit is $Z_\mathrm{gauss} / Z_\mathrm{power} = 69 \pm 1.1 \e{7}$ indicating a very strong preference for the gaussian model. We also compare to the mass ratio distribution expected for random pairing from the Kroupa IMF. In this case, there are no free parameters so the evidence integral just becomes the likelihood function (Equation \ref{eqn:money}). The corresponding odds ratio is  $Z_\mathrm{gauss} / Z_\mathrm{kroupa} = 333 \pm 4 \e{17}$, again indicating a very strong preference for the gaussian fit.

TODO: Speculate on the discrepancy between my results and VAST. Perhaps the change occurs closer than their cut of 125 AU? Perhaps it is inside $\sim 30$ AU, where they have very low completeness?

>>>>

\begin{figure}
\includegraphics[width=\columnwidth]{Imaging_Data.pdf}
\caption{Detection images for all stars in which we detect a companion in the follow-up NIRI data. 
TODO: Update imaging table. What to do with HIP 88116???}
\label{fig:images}
\end{figure}

@@

\begin{equation}
Q(T) = \sum_k Q(T, v_k) v_k 
\end{equation} 
where $Q(T, v\sin{i})$ is the surface plotted in Figure \ref{fig:detrate_2d} and $v_k$ are the samples from the distribution of $v\sin{i}$. For $T < 6000$ K, we sample $v\sin{i}$ using the gyrochronology relation given in \citet{Barnes2010b}:

@@

We show the marginalized detection rate and mean value of $v\sin{i}$ as a function of temperature in Figure \ref{fig:marginalized}. Both the detection rate and the average $v\sin{i}$ are smoothly varying, and show the expected behaviour with temperature. The detection rate falls with hotter temperatures because the companions are expected to be fast rotators, which are more difficult to detect.

@@

@@

In the above equation, $\{\vec{x_k}\}$ denotes the full spectrum of star $k$, and $\vec{\theta}$ denotes the parameters for the model we are fitting. $K=360$ is the total sample size for our survey, $Q(q)$ is the completeness function shown in Figure \ref{fig:completeness}, and $P(q|\vec{\theta})$ is the likelihood function for the mass ratio given the model parameters. We fit the data to three distinct distributions: a histogram ($P_1$) and a beta distribution ($P_2$), and a power law ($P_3$):

\begin{align}
 P_1(q|\vec{\theta}) &= \begin{cases}
      \hfill \theta_1 \hfill & q \in \Delta_1 \\
      \hfill \theta_2 \hfill & q \in \Delta_2 \\
      \hfill \ldots \\
      \hfill \theta_7 \hfill & q \in \Delta_7
     \end{cases} \label{eqn:P1} \\
 %P_2(q|\vec{\theta}) &= \frac{A}{\sqrt{2\pi\sigma^2}} e^{-\frac{(q - \mu)^2}{2\sigma^2}} \label{eqn:P2} \\
 P_2(q|\vec{\theta}) &= \frac{\Gamma(a+b)}{\Gamma(a)\Gamma(b)} q^{a-1}(1-q)^{b-1} \\
 P_3(q|\vec{\theta}) &= (1-\gamma)q^{-\gamma} \label{eqn:P3}
\end{align}
%The constant $A$ in the gaussian distribution is a renormalization factor such that the distribution is only defined from $0 < q < 1$:

%\begin{equation}
%A = \frac{2}{\erf\left(\frac{\mu}{\sigma \sqrt{2}}\right) - \erf\left(\frac{\mu - 1}{\sigma \sqrt{2}}\right)}
%\end{equation}

@@

The beta function fit only has two parameters ($a, b$), and was chosen because it has a similar shape to the histogram resulting from the first model and is only defined for $0 < q < 1$. We use uniform priors on both $a$ and $b$, resulting in the posterior probability

\begin{equation}
P_2(a, b | \{\vec{x_k}\}) \propto P_2(q|a, b) 
\end{equation}

@@

We show the resulting fits in Figure \ref{fig:mrd}. The $1\sigma$ uncertainties in the bin heights from the histogram model are shown as error bars, and we overplot 500 samples of the gaussian fit to show the spread allowed by the data. The best-fit power law is plotted with a green dot-dashed line, as well as the mass ratio distribution expected from random pairing from the Kroupa initial mass function (IMF) in yellow.

@@

\begin{figure}
\includegraphics[width=\columnwidth]{MRD_total.pdf}
\caption{Mass ratio distribution for our sample. The data was fit to a histogram, a beta distribution, and a power law. The histogram is shown in the solid blue blocks, with $1 \sigma$ uncertainties marked with error bars. The variance of the beta function fit is shown with 500 samples from the posterior probability distribution for the parameters. We also show the best-fit power law and the mass ratio distribution resulting from random pairing of the Kroupa initial mass function.}
\label{fig:mrd}
\end{figure}

@@

The most striking feature of the mass ratio distribution shown in Figure \ref{fig:mrd} is the turnover at intermediate q, which the beta distribution fit finds occurs at $q = 0.30^{+0.03}_{-0.04}$. Although the power law fit and the Kroupa IMF are visually very poor fits to the data, we formally compare the models to ensure that the different form is not an artifact of binning or simple noise. We make the comparison using the posterior odds:

@@

The integrals are estimated as part of the nested sampling algorithm in the MultiNest code. The odds ratio comparing the beta distribution to power law fit is $Z_\mathrm{beta} / Z_\mathrm{power} = 81 \pm 1.3 \e{8}$ indicating a very strong preference for the beta distribution model. We also compare to the mass ratio distribution expected for random pairing from the Kroupa IMF. In this case, there are no free parameters so the evidence integral just becomes the likelihood function (Equation \ref{eqn:money}). The corresponding odds ratio is  $Z_\mathrm{beta} / Z_\mathrm{kroupa} = 325 \pm 3.6 \e{18}$, again indicating a very strong preference for the beta distribution fit.

\subsection{Comparison to Previous Results}

Our mass ratio distribution appears to be in tension with the results of the VAST survey \citep{DeRosa2014}, which finds a nearly flat distribution for close companions. However, their subsample of close companions only includes 18 binaries, so it is possible that the different forms are just a result of small number statistics. To assess the degree of tension, we use the Anderson-Darling [CITE] test to find the probability that both their close companion subsample and our companions are drawn from the same parent distribution. We only use companions from this work with mean $\overline{q} > 0.15$, since the survey completeness is roughly constant in that range. To account for measurement uncertainties in the mass ratios, we choose from the mass ratio samples (see Section \ref{sec:mrd}) many times and compute the Anderson-Darling test statistic each time. Since \citet{DeRosa2014} do not quote uncertainties, we assume uncertainties of $\sigma_q = 0.05$ for all of their measurements. The result is $p = 0.02^{+0.02}_{-0.01}$, indicating that it is unlikely that both samples come from the same parent distribution.

The differing results indicate that the mass ratio distribution changes inside $\sim 30-50$ AU, where \citet{DeRosa2014} has low completeness. This intrepretation implies that the mass ratio distribution is a very steep power law at wide ($a \gtrsim 150$ AU) separations, flattens out at intermediate separations ($50\ \mathrm{AU} \lesssim a \lesssim 150\ \mathrm{AU}$), and begins to show a preference for $q \sim 0.3$ at close separations ($a \lesssim 50$ AU).

\subsection{Theoretical Implications}
